---
title: "Choosing a Classification Method for Measuring Government Criticism in Armenian Civil Society Texts"
author: "Albert Ananyan"
date: "2026-03-16"
format:
  pdf:
    toc: true
    number-sections: true
    geometry: margin=1in
    fontsize: 12pt
    linestretch: 1.15
execute:
  warning: false
  message: false
bibliography: references.bib
---

```{r}
library(tidyverse)
library(reticulate)
library(knitr)
library(kableExtra)

opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Data Loading and Preparation

## Load the coded training corpus

Loading the combined training corpus with 642 hand-coded documents from five Armenian watchdog CSOs. Each document is coded for:

- `crit_armenian_human`: 1 if the document criticizes the Armenian government
- `crit_foreign_human`: 1 if the document criticizes foreign actors

```{r}
coded_df <- read_csv("data/coded_corpus_combined.csv", show_col_types = FALSE)

cat("Total coded documents:", nrow(coded_df), "\n\n")

org_summary <- coded_df %>%
  group_by(Organization) %>%
  summarise(
    n = n(),
    critical = sum(crit_armenian_human == 1),
    non_critical = sum(crit_armenian_human == 0),
    pct_critical = round(mean(crit_armenian_human) * 100, 1)
  )

kable(org_summary,
      caption = "Training data distribution by organization",
      col.names = c("Organization", "N", "Critical", "Non-critical", "\\% Critical"))
```

## Visualize class distribution

```{r}
coded_df %>%
  mutate(
    label = ifelse(crit_armenian_human == 1, "Critical", "Non-critical"),
    label = factor(label, levels = c("Critical", "Non-critical"))
  ) %>%
  count(Organization, label) %>%
  ggplot(aes(x = reorder(Organization, -n), y = n, fill = label)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Critical" = "#E74C3C", "Non-critical" = "#3498DB")) +
  labs(
    title = "Training Set Composition by Organization",
    x = NULL, y = "Number of documents", fill = "Label"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Filter foreign-critical documents

Excluding documents that only criticize foreign actors so the classifier specifically captures government-directed criticism.

```{r}
train_df <- coded_df %>%
  filter(crit_foreign_human == 0 | crit_armenian_human == 1) %>%
  transmute(
    Organization = Organization,
    text = full_text,
    label = as.integer(crit_armenian_human)
  ) %>%
  drop_na(text, label)

cat("Documents after filtering:", nrow(train_df), "\n")
cat("Critical:", sum(train_df$label == 1),
    "| Non-critical:", sum(train_df$label == 0), "\n")
cat("Class ratio:", round(sum(train_df$label == 1) / sum(train_df$label == 0), 2), ": 1\n")
```

# Method 1: Naive Bayes on Bag-of-Words

## Preprocessing

```{r}
library(quanteda)
library(quanteda.textmodels)
library(quanteda.textstats)

# custom Armenian stopwords
armenian_stops <- jsonlite::fromJSON("data/stopwords-hy-clean.json")
cat("Loaded", length(armenian_stops), "Armenian stopwords\n")
```

```{r}
corpus_cso <- corpus(train_df, text_field = "text")

# tokenize and remove stopwords
toks <- corpus_cso %>%
  tokens(remove_punct = TRUE,
         remove_symbols = TRUE,
         remove_numbers = TRUE) %>%
  tokens_remove(armenian_stops)

dfm_cso <- dfm(toks)
cat("Initial features:", nfeat(dfm_cso), "\n")

# trim rare words
dfm_cso <- dfm_trim(dfm_cso, min_docfreq = 2)
cat("After trimming:", nfeat(dfm_cso), "features\n")
```

## Cross-validation

5-fold stratified CV, same folds used across all models.

```{r}
set.seed(123)

valid_docs <- !is.na(docvars(dfm_cso)$label)
dfm_valid <- dfm_subset(dfm_cso, valid_docs)

k <- 5
folds <- sample(rep(1:k, length.out = ndoc(dfm_valid)))

cv_results_nb <- data.frame(fold = 1:k, accuracy = NA, f1 = NA,
                             precision = NA, recall = NA)

for(i in 1:k) {
  dfm_cv_train <- dfm_subset(dfm_valid, folds != i)
  dfm_cv_test <- dfm_subset(dfm_valid, folds == i)

  train_labels <- factor(docvars(dfm_cv_train)$label,
                         levels = c(0, 1),
                         labels = c("non_critical", "critical"))

  model_cv <- textmodel_nb(dfm_cv_train, train_labels)

  dfm_cv_test_matched <- dfm_match(dfm_cv_test, features = featnames(dfm_cv_train))
  pred <- predict(model_cv, newdata = dfm_cv_test_matched)
  actual <- factor(docvars(dfm_cv_test_matched)$label,
                   levels = c(0, 1),
                   labels = c("non_critical", "critical"))

  cv_results_nb$accuracy[i] <- mean(pred == actual, na.rm = TRUE)

  tp <- sum(pred == "critical" & actual == "critical")
  fp <- sum(pred == "critical" & actual == "non_critical")
  fn <- sum(pred == "non_critical" & actual == "critical")

  prec <- ifelse(tp + fp > 0, tp / (tp + fp), 0)
  rec <- ifelse(tp + fn > 0, tp / (tp + fn), 0)

  cv_results_nb$precision[i] <- prec
  cv_results_nb$recall[i] <- rec
  cv_results_nb$f1[i] <- ifelse(prec + rec > 0, 2 * prec * rec / (prec + rec), 0)
}

cat("=== Naive Bayes (Bag-of-Words) ===\n")
cat("Accuracy:", sprintf("%.1f%%", mean(cv_results_nb$accuracy) * 100),
    "+/-", sprintf("%.1f%%", sd(cv_results_nb$accuracy) * 100), "\n")
cat("F1 (critical):", sprintf("%.3f", mean(cv_results_nb$f1)),
    "+/-", sprintf("%.3f", sd(cv_results_nb$f1)), "\n")
cat("Precision:", sprintf("%.3f", mean(cv_results_nb$precision)), "\n")
cat("Recall:", sprintf("%.3f", mean(cv_results_nb$recall)), "\n")
```

## Fold-level accuracy

```{r}
ggplot(cv_results_nb, aes(x = factor(fold), y = accuracy)) +
  geom_col(fill = "#95A5A6", width = 0.6) +
  geom_hline(yintercept = mean(cv_results_nb$accuracy),
             linetype = "dashed", color = "#E74C3C") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(title = "Naive Bayes: Accuracy by Fold",
       x = "Fold", y = "Accuracy") +
  theme_minimal()
```

# Methods 2-5: BERT Embeddings + Classifiers

Using HyeBERT (Armenian BERT) to get 768-dim [CLS] embeddings as document representations, then training classifiers on top.

## Generate BERT embeddings

```{python}
import os, warnings
import numpy as np
import pandas as pd
import torch
from transformers import AutoTokenizer, AutoModel

warnings.filterwarnings("ignore")

model_name = "aking11/hyebert"
print(f"Loading model: {model_name}")
tokenizer = AutoTokenizer.from_pretrained(model_name)
bert_model = AutoModel.from_pretrained(model_name)
bert_model.eval()

def get_embedding(text, tokenizer, model, max_length=512):
    """Extract [CLS] token embedding from BERT."""
    inputs = tokenizer(str(text), return_tensors="pt",
                       truncation=True, max_length=max_length, padding=True)
    with torch.no_grad():
        outputs = model(**inputs)
    return outputs.last_hidden_state[:, 0, :].cpu().numpy().flatten()

def embed_series(texts, tokenizer, model, step=50):
    """Embed all documents, printing progress."""
    vecs = []
    for i, t in enumerate(texts, start=1):
        if i == 1 or i % step == 0 or i == len(texts):
            print(f"  Embedding {i}/{len(texts)}")
        vecs.append(get_embedding(t, tokenizer, model))
    return np.array(vecs)

train_df = r.train_df
X = embed_series(train_df["text"], tokenizer, bert_model)
y = train_df["label"].astype(int).values

print(f"\nEmbedding matrix: {X.shape[0]} docs x {X.shape[1]} dimensions")
```

## Train and evaluate classifiers

Training four classifiers on frozen BERT embeddings using the same 5-fold CV. Not using class_weight="balanced" since training data is already roughly balanced.

```{python}
from sklearn.linear_model import LogisticRegression, LogisticRegressionCV
from sklearn.ensemble import RandomForestClassifier, GradientBoostingClassifier
from sklearn.model_selection import StratifiedKFold, cross_val_score, cross_val_predict
from sklearn.metrics import classification_report, confusion_matrix

cv = StratifiedKFold(n_splits=5, shuffle=True, random_state=42)

lr = LogisticRegression(max_iter=2000, random_state=42)

rf = RandomForestClassifier(n_estimators=500, min_samples_leaf=3,
                            random_state=42, n_jobs=-1)

gb = GradientBoostingClassifier(n_estimators=300, max_depth=4, learning_rate=0.05,
                                subsample=0.8, random_state=42)

results = {}
for name, model in [("Logistic Regression", lr),
                     ("Random Forest", rf),
                     ("Gradient Boosting", gb)]:
    acc = cross_val_score(model, X, y, cv=cv, scoring="accuracy")
    f1 = cross_val_score(model, X, y, cv=cv, scoring="f1")
    prec = cross_val_score(model, X, y, cv=cv, scoring="precision")
    rec = cross_val_score(model, X, y, cv=cv, scoring="recall")
    pred = cross_val_predict(model, X, y, cv=cv)
    results[name] = {"acc": acc, "f1": f1, "prec": prec, "rec": rec, "pred": pred}
    print(f"{name}:")
    print(f"  Accuracy  = {acc.mean():.3f} +/- {acc.std():.3f}")
    print(f"  F1        = {f1.mean():.3f} +/- {f1.std():.3f}")
    print(f"  Precision = {prec.mean():.3f}")
    print(f"  Recall    = {rec.mean():.3f}\n")

# Stacking: cross-validated probabilities from each base model -> meta-learner
lr_p = cross_val_predict(lr, X, y, cv=cv, method="predict_proba")[:, 1]
rf_p = cross_val_predict(rf, X, y, cv=cv, method="predict_proba")[:, 1]
gb_p = cross_val_predict(gb, X, y, cv=cv, method="predict_proba")[:, 1]

X_stack = np.column_stack([lr_p, rf_p, gb_p])

meta = LogisticRegressionCV(cv=5, random_state=42, max_iter=2000)
stack_acc = cross_val_score(meta, X_stack, y, cv=cv, scoring="accuracy")
stack_f1 = cross_val_score(meta, X_stack, y, cv=cv, scoring="f1")
stack_prec = cross_val_score(meta, X_stack, y, cv=cv, scoring="precision")
stack_rec = cross_val_score(meta, X_stack, y, cv=cv, scoring="recall")
stack_pred = cross_val_predict(meta, X_stack, y, cv=cv)
results["Stacking Ensemble"] = {"acc": stack_acc, "f1": stack_f1,
                                  "prec": stack_prec, "rec": stack_rec,
                                  "pred": stack_pred}

print(f"Stacking Ensemble:")
print(f"  Accuracy  = {stack_acc.mean():.3f} +/- {stack_acc.std():.3f}")
print(f"  F1        = {stack_f1.mean():.3f} +/- {stack_f1.std():.3f}")
print(f"  Precision = {stack_prec.mean():.3f}")
print(f"  Recall    = {stack_rec.mean():.3f}")
```

## Summary table

```{python}
# comparing all five methods
nb_acc = float(r.cv_results_nb['accuracy'].mean())
nb_f1 = float(r.cv_results_nb['f1'].mean())
nb_prec = float(r.cv_results_nb['precision'].mean())
nb_rec = float(r.cv_results_nb['recall'].mean())
nb_acc_sd = float(r.cv_results_nb['accuracy'].std())

rows = []
rows.append(("Naive Bayes", "Bag-of-words", nb_acc, nb_acc_sd, nb_f1, nb_prec, nb_rec))
for name in ["Logistic Regression", "Random Forest", "Gradient Boosting", "Stacking Ensemble"]:
    r_dict = results[name]
    rows.append((name, "BERT", r_dict['acc'].mean(), r_dict['acc'].std(),
                 r_dict['f1'].mean(), r_dict['prec'].mean(), r_dict['rec'].mean()))

print(f"\n{'Method':<22} {'Repr.':<8} {'Acc':>7} {'SD':>6} {'F1':>6} {'Prec':>6} {'Rec':>6}")
print("-" * 65)
for row in rows:
    print(f"{row[0]:<22} {row[1]:<8} {row[2]:>6.1%} {row[3]:>6.3f} {row[4]:>6.3f} {row[5]:>6.3f} {row[6]:>6.3f}")
```

## Fine-tuning vs. feature extraction

Testing whether fine-tuning HyeBERT's last 2 transformer layers outperforms frozen embeddings.

```{python}
from transformers import AutoModelForSequenceClassification
from torch.utils.data import Dataset as TorchDataset, DataLoader

BATCH_SIZE = 8
EPOCHS = 5
FT_LR = 2e-5
FREEZE_LAYERS = 10
device = torch.device("mps" if torch.backends.mps.is_available() else "cpu")

class CSODataset(TorchDataset):
    def __init__(self, texts, labels, tok, max_len=512):
        self.enc = tok(texts, truncation=True, padding=True,
                       max_length=max_len, return_tensors="pt")
        self.labels = torch.tensor(labels, dtype=torch.long)
    def __len__(self):
        return len(self.labels)
    def __getitem__(self, idx):
        return {"input_ids": self.enc["input_ids"][idx],
                "attention_mask": self.enc["attention_mask"][idx],
                "labels": self.labels[idx]}

texts_list = train_df["text"].astype(str).tolist()
labels_arr = train_df["label"].astype(int).values

ft_preds = np.zeros(len(labels_arr))
fe_preds = np.zeros(len(labels_arr))

for fold, (tr_idx, va_idx) in enumerate(cv.split(texts_list, labels_arr)):
    print(f"Fold {fold+1}/5")
    tr_texts = [texts_list[i] for i in tr_idx]
    va_texts = [texts_list[i] for i in va_idx]
    tr_labels = labels_arr[tr_idx]
    va_labels = labels_arr[va_idx]

    # fine-tuned model
    ft_model = AutoModelForSequenceClassification.from_pretrained(
        model_name, num_labels=2)
    for pname, param in ft_model.named_parameters():
        if "encoder.layer" in pname:
            layer_num = int(pname.split("encoder.layer.")[1].split(".")[0])
            if layer_num < FREEZE_LAYERS:
                param.requires_grad = False
        elif "embeddings" in pname:
            param.requires_grad = False
    ft_model.to(device)

    tr_loader = DataLoader(CSODataset(tr_texts, tr_labels, tokenizer),
                           batch_size=BATCH_SIZE, shuffle=True)
    va_loader = DataLoader(CSODataset(va_texts, va_labels, tokenizer),
                           batch_size=BATCH_SIZE)
    opt = torch.optim.AdamW(
        [p for p in ft_model.parameters() if p.requires_grad],
        lr=FT_LR, weight_decay=0.01)

    for epoch in range(EPOCHS):
        ft_model.train()
        for batch in tr_loader:
            opt.zero_grad()
            out = ft_model(input_ids=batch["input_ids"].to(device),
                           attention_mask=batch["attention_mask"].to(device),
                           labels=batch["labels"].to(device))
            out.loss.backward()
            torch.nn.utils.clip_grad_norm_(ft_model.parameters(), 1.0)
            opt.step()

    ft_model.eval()
    fold_p = []
    with torch.no_grad():
        for batch in va_loader:
            out = ft_model(input_ids=batch["input_ids"].to(device),
                           attention_mask=batch["attention_mask"].to(device))
            fold_p.extend(torch.argmax(out.logits, dim=-1).cpu().numpy())
    ft_preds[va_idx] = fold_p

    # frozen embeddings baseline
    from sklearn.linear_model import LogisticRegression as LR_
    lr_bl = LR_(max_iter=2000, random_state=42)
    lr_bl.fit(X[tr_idx], tr_labels)
    fe_preds[va_idx] = lr_bl.predict(X[va_idx])

    del ft_model
    if device.type == "mps":
        torch.mps.empty_cache()

from sklearn.metrics import accuracy_score, f1_score, precision_score, recall_score

print(f"\n{'Method':<25} {'Acc':>7} {'F1':>6} {'Prec':>6} {'Rec':>6}")
print("-" * 55)
for lbl, p in [("Fine-tuned HyeBERT", ft_preds),
               ("Frozen + LogReg", fe_preds)]:
    print(f"{lbl:<25} {accuracy_score(labels_arr, p):>6.1%} "
          f"{f1_score(labels_arr, p):>6.3f} "
          f"{precision_score(labels_arr, p):>6.3f} "
          f"{recall_score(labels_arr, p):>6.3f}")

delta_acc = (accuracy_score(labels_arr, ft_preds) -
             accuracy_score(labels_arr, fe_preds)) * 100
delta_f1  = (f1_score(labels_arr, ft_preds) -
             f1_score(labels_arr, fe_preds)) * 100
print(f"\nDelta: Accuracy {delta_acc:+.1f}pp, F1 {delta_f1:+.1f}pp")
```

## Performance comparison

```{r}
perf_df <- tibble(
  method = c("Naive Bayes", "Logistic Reg.", "Random Forest",
             "Gradient Boost.", "Stacking Ens."),
  representation = c("Bag-of-words", rep("BERT", 4)),
  accuracy = c(
    mean(cv_results_nb$accuracy),
    mean(py$results[["Logistic Regression"]][["acc"]]),
    mean(py$results[["Random Forest"]][["acc"]]),
    mean(py$results[["Gradient Boosting"]][["acc"]]),
    mean(py$results[["Stacking Ensemble"]][["acc"]])
  ),
  sd = c(
    sd(cv_results_nb$accuracy),
    sd(py$results[["Logistic Regression"]][["acc"]]),
    sd(py$results[["Random Forest"]][["acc"]]),
    sd(py$results[["Gradient Boosting"]][["acc"]]),
    sd(py$results[["Stacking Ensemble"]][["acc"]])
  )
)

perf_df$method <- factor(perf_df$method, levels = perf_df$method)

ggplot(perf_df, aes(x = method, y = accuracy, fill = representation)) +
  geom_col(width = 0.65) +
  geom_errorbar(aes(ymin = accuracy - sd, ymax = accuracy + sd), width = 0.2) +
  geom_vline(xintercept = 1.5, linetype = "dashed", color = "gray40") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_fill_manual(values = c("Bag-of-words" = "#95A5A6", "BERT" = "#2C3E50")) +
  labs(
    title = "Classification Accuracy Across Methods",
    subtitle = "Error bars show +/- 1 SD across 5 CV folds",
    x = NULL, y = "Accuracy", fill = "Representation"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Model agreement

```{python}
# checking how often BERT-based classifiers agree with each other
pred_df = pd.DataFrame({
    "LR": results["Logistic Regression"]["pred"],
    "RF": results["Random Forest"]["pred"],
    "GB": results["Gradient Boosting"]["pred"],
    "Stack": results["Stacking Ensemble"]["pred"]
})

pairs = [("LR","RF"), ("LR","GB"), ("RF","GB")]
print("Pairwise agreement:\n")
for a, b in pairs:
    print(f"  {a} vs {b}: {(pred_df[a] == pred_df[b]).mean():.1%}")

all_agree = (pred_df["LR"] == pred_df["RF"]) & (pred_df["RF"] == pred_df["GB"])
print(f"\nAll three agree: {all_agree.mean():.1%} of documents")
```

```{r}
agree_rate <- mean(py$all_agree)
agree_df <- tibble(
  status = c("All agree", "Disagree"),
  pct = c(agree_rate, 1 - agree_rate)
)

ggplot(agree_df, aes(x = "", y = pct, fill = status)) +
  geom_col(width = 0.4) +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = c("All agree" = "#27AE60", "Disagree" = "#E74C3C")) +
  labs(title = "Model Agreement Rate", x = NULL, y = NULL, fill = NULL) +
  theme_minimal()
```

# Full Corpus Classification

## Load full corpus

```{r}
uic <- read_csv("data/uic_publications.csv", show_col_types = FALSE)
hcav <- read_csv("data/hcav_publications.csv", show_col_types = FALSE)
hahr <- read_csv("data/hahr_publications.csv", show_col_types = FALSE)
csi <- read_csv("data/csi_publications.csv", show_col_types = FALSE)
ti <- read_csv("data/ti_armenia_publications.csv", show_col_types = FALSE)

harmonize <- function(df, org_name) {
  df %>%
    mutate(Organization = org_name,
           full_text = paste(title, content, sep = " ")) %>%
    filter(!is.na(content), nchar(content) > 100) %>%
    select(Organization, title, date, full_text)
}

full_corpus <- bind_rows(
  harmonize(uic, "UIC"),
  harmonize(hcav, "HCAV"),
  harmonize(hahr, "HAHR"),
  harmonize(csi, "CSI"),
  harmonize(ti, "TI_ARMENIA")
)

cat("Full corpus:", nrow(full_corpus), "documents\n\n")
full_corpus %>%
  group_by(Organization) %>%
  summarise(n = n()) %>%
  kable(caption = "Documents by organization in full corpus")
```

## Classify with Random Forest

Using RF since it had the highest accuracy in cross-validation.

```{python}
rf_final = RandomForestClassifier(n_estimators=500, min_samples_leaf=3,
                                   random_state=42, n_jobs=-1)
rf_final.fit(X, y)

full_df = r.full_corpus
print(f"Embedding full corpus: {len(full_df)} documents...")
X_full = embed_series(full_df["full_text"], tokenizer, bert_model, step=500)

full_preds = rf_final.predict(X_full)
full_probs = rf_final.predict_proba(X_full)

full_df["predicted_class"] = ["critical" if p == 1 else "non_critical" for p in full_preds]
full_df["prob_critical"] = full_probs[:, 1]

print(f"\nPredicted critical: {sum(full_preds)} ({sum(full_preds)/len(full_preds):.1%})")
print(f"Predicted non-critical: {sum(full_preds==0)} ({sum(full_preds==0)/len(full_preds):.1%})")
```

## Criticality trajectories

```{r}
classified <- py$full_df %>%
  mutate(year = lubridate::year(as.Date(date)))

cso_year_panel <- classified %>%
  filter(!is.na(year), year >= 2014, year <= 2024) %>%
  group_by(Organization, year) %>%
  summarise(
    n_docs = n(),
    n_critical = sum(predicted_class == "critical"),
    critical_share = mean(predicted_class == "critical"),
    .groups = "drop"
  )

ggplot(cso_year_panel, aes(x = year, y = critical_share,
                            color = Organization, group = Organization)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_vline(xintercept = 2018, linetype = "dashed", color = "gray50") +
  annotate("text", x = 2018.2, y = 0.05, label = "Velvet Revolution",
           hjust = 0, size = 3, color = "gray40") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  scale_x_continuous(breaks = 2014:2024) +
  labs(
    title = "Figure 6. Predicted Criticality by CSO-Year",
    x = "Year", y = "Critical Share"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

write_csv(cso_year_panel, "data/cso_year_panel_2014_2024.csv")
```

## Document volume

```{r}
ggplot(cso_year_panel, aes(x = year, y = n_docs, fill = Organization)) +
  geom_col() +
  facet_wrap(~Organization, scales = "free_y", ncol = 1) +
  scale_x_continuous(breaks = 2014:2024) +
  labs(
    title = "Document Volume by CSO-Year",
    x = "Year", y = "Number of documents"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))
```

# Sensitivity Analysis

Checking whether different classifiers produce different substantive conclusions.

```{python}
lr_final = LogisticRegression(max_iter=2000, random_state=42)
gb_final = GradientBoostingClassifier(n_estimators=300, max_depth=4, learning_rate=0.05,
                                       subsample=0.8, random_state=42)
lr_final.fit(X, y)
gb_final.fit(X, y)

lr_preds = lr_final.predict(X_full)
gb_preds = gb_final.predict(X_full)
rf_preds = full_preds

full_df["pred_lr"] = lr_preds
full_df["pred_rf"] = rf_preds
full_df["pred_gb"] = gb_preds

agree_all = ((lr_preds == rf_preds) & (rf_preds == gb_preds)).mean()
print(f"All three classifiers agree on {agree_all:.1%} of full corpus documents")
```

```{r}
sensitivity_df <- py$full_df %>%
  mutate(year = lubridate::year(as.Date(date))) %>%
  filter(!is.na(year), year >= 2014, year <= 2024) %>%
  group_by(Organization, year) %>%
  summarise(
    LR = mean(pred_lr == 1),
    RF = mean(pred_rf == 1),
    GB = mean(pred_gb == 1),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(LR, RF, GB),
               names_to = "Classifier",
               values_to = "critical_share")

ggplot(sensitivity_df, aes(x = year, y = critical_share,
                            color = Classifier, linetype = Classifier)) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 1.5) +
  facet_wrap(~Organization) +
  geom_vline(xintercept = 2018, linetype = "dashed", color = "gray70", linewidth = 0.3) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(breaks = seq(2014, 2024, 2)) +
  labs(
    title = "Sensitivity Analysis: Criticality Trajectories by Classifier",
    x = "Year", y = "Share of critical publications"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Document-Level Logistic Regression

Estimating a document-level logit with organization fixed effects and cluster-robust SEs to formally test whether criticism changed after 2018.

```{python}
import statsmodels.formula.api as smf
from statsmodels.stats.sandwich_covariance import cov_cluster
from scipy import stats as sp_stats

doc_df = full_df.copy()
doc_df['critical_bin'] = (doc_df['predicted_class'] == 'critical').astype(int)
doc_df['year_num'] = pd.to_datetime(doc_df['date']).dt.year
doc_df = doc_df[(doc_df['year_num'] >= 2014) & (doc_df['year_num'] <= 2024)].copy()
doc_df['post_2018'] = (doc_df['year_num'] >= 2019).astype(int)
doc_df['org_year'] = doc_df['Organization'] + '_' + doc_df['year_num'].astype(str)

print(f"Documents: {len(doc_df)}")
print(f"Pre-2018 criticality: {doc_df[doc_df['post_2018']==0]['critical_bin'].mean():.1%}")
print(f"Post-2018 criticality: {doc_df[doc_df['post_2018']==1]['critical_bin'].mean():.1%}")

model1 = smf.logit('critical_bin ~ post_2018 + C(Organization)', data=doc_df).fit(disp=0)

clust_cov = cov_cluster(model1, doc_df['org_year'])
clust_se = np.sqrt(np.diag(clust_cov))
z_vals = model1.params / clust_se
p_vals = 2 * (1 - sp_stats.norm.cdf(np.abs(z_vals)))

print("\n--- Logit with Org Fixed Effects ---")
print(f"{'Variable':<35} {'Coef':>8} {'SE(cl)':>8} {'z':>7} {'p':>7}")
print("-" * 70)
for name, coef, se, z, p in zip(model1.params.index, model1.params, clust_se, z_vals, p_vals):
    sig = '***' if p < 0.001 else '**' if p < 0.01 else '*' if p < 0.05 else ''
    print(f"{name:<35} {coef:>8.4f} {se:>8.4f} {z:>7.2f} {p:>7.4f} {sig}")

# average marginal effect
p_hat = model1.predict(doc_df)
beta = model1.params['post_2018']
ame = (p_hat * (1 - p_hat) * beta).mean()
print(f"\nAverage Marginal Effect of Post-2018: {ame:.4f}")
print(f"Interpretation: {abs(ame)*100:.1f} pp {'decrease' if ame < 0 else 'increase'} "
      f"in probability of critical content (p = {p_vals[list(model1.params.index).index('post_2018')]:.3f})")
```

# Session Info

```{r}
sessionInfo()
```
